---
author:
  - name: Matthieu Stigler
    affiliation: UC Davis
    email: matthieu.stigler@gmail.com
    address: California
    url: https://matthieustigler.github.io/
  - name: Bastiaan Quast
    affiliation: The Graduate Institute, Geneva
    address: >
      Maison de la paix
      Geneva, Switzerland
    email: bquast@gmail.com
    url: http://qua.st/
title:
  formatted: "\\pkg{rddtools}: tools for Regression Discontinuity Design in R\\thanks{This research was financed by the Swiss National Science Foundation (SNSF) under the grant 'Development Aid and Social Dynamics' (100018-140745) administered by the Graduate Institute's Centre on Conflict, Development and Peacebuilding (CCDP) and led by Jean-Louis Arcand. We thank Sandra Reimann and Oliver Jutersonke at the CCDP for their generous support.}"
  plain:     "rddtools: tools for Regression Discontinuity Design in R"
  short:     "\\pkg{rddtools}"
abstract: >
  The rddtools package provides a framework for Regression Discontinuity Design (RDD) in R.
  In addtion to bringing together functionality from several different existing package,
  new functionality is implemented in terms of design and sensitivity test,
  as well as non parametric RDD.
keywords:
  formatted: [RDD, Regression, Discontinuity, Design, "\\proglang{R}"]
  plain:     [RDD, Regression, Discontinuity, Design R]
submitdate: 2015-07-25  
preamble: >
  \usepackage{amsmath}
output:
  md_document
   
  rticles::jss_article 
bibliography: bibliography.bib
---

# Introduction

The `rddtools` package attempts to provide a unified approach to the application of Regression Discontinuity Design (RDD) in R.
Functionality from several existing packages is brought together under one coherent API.
Additionallity, the `rddtools` package implements new functionality in several aspects of regression discontinuity design.


# Design

A unified framework for RDD is implemented through the `rdd_data` class.
This class inherits from the `R` `base` package's `data.frame` class.
This functionality is made accessible through the associated `rdd_data()` function,
as well as the following associated methods.

- `[.rdd_data()` / `subset.rdd_data()`
- `summary.rdd_data()`
- `plot.rdd_data()`

The package is designed to leveredge of existing implementations of **Regression Discontinuity Design** in `R`, such as the `rdd` [@dimmery2013rdd] and `KernSmooth` [@wand2015kernsmooth] packages.
Furthermore, general algorithms such as non-parametric regression from the `np` package [@hayfield2008np] is made accessible for RDD through the `rdd_data` framework.

In addition to this, it implements several tools for RDD analysis that were previously unavailable.


## Bandwidth Selection

The `rddtools` package implements two new methods for bandwidth selection.
The first is the MSE-RDD bandwidth procedure of [@imbens2012optimal].
This procdure is implemented in the `rdd_bw_ik()` function. 
Secondly, the MSE global bandwidth procedure of [@ruppert1995effective] is implemented in the `rdd_bw_rsw()` function.
    
    
## Estimation

Various types of RDD estimation are supported.
The functionality has been implemented in such a way,
that the change from one estimation method to another is as small as possible.

Firstly, RDD parametric estimation though the `rdd_reg_lm()` function is implemented.
The `rdd-reg_lm()` fnction includes functionality for specifying the polynomial order,
including covariates with various specifications as advocated in [@imbens2008regression].

Secondly, RDD local non-parametric estimation is available by means of the `rdd_reg_np()` function.
The `rdd_reg_np()` function can also include covariates
and allows different types of inference ( such as fully non-parametric, or parametric approximation).

Lastly, RDD generalised estimation has been implemented.
This allows to use custom estimating functions to get the RDD coefficient.
For example a probit RDD, or quantile regression could be used here.
    

## Post-Estimation
    
A collection of Post-Estimation tools allow the robustness of the estimation results to be verified.

This includes various tools, such as the `rdd_pred()`, which is used to obtain predictions at given covariate values.
As well as the `as.lm()` function, which is used to convert to the lm class.
Furthermore there is the `as.npreg()` function, in order to convert to the `np` package.

Additional post-estmimation tools include `clusterInf()`, which can be used for inference with clustered data,
using eithera covariance matrix (using the `vcovCluster()` function),
or by a degrees of freedom corrention (as described in [@cameron2008bootstrap]).

Finally, the package contains functions to replicate the Monte-Carlo simulations of [Imbens and Kalyanaraman 2012],
using the `gen_mc_ik()` function.


## Regression Sensitivity Analysis

Regression sensitivity analysis can be conducted using the `plotSense()` function,
which test the sensitivity of the coefficient with respect to the bandwidth,
or by means of **Placebo plot** using different cutpoints: `plotPlacebo()`


## Design sensitivity analysis

Finally, methods for design sensitivity analysis are included.

The McCrary test of manipulation of the forcing variable is available by passing the wrapper `dens_test()` to the function `DCdensity()` from package `rdd`. 

As well as, the test of equal means of covariates, which can be performed using the `covarTest_mean()` function.

In addition to this, the test of equal density of covariates is available via the `covarTest_dens()` function.


# Data

A collection of typical data sets is included in the package.

- Initiative Nationale pour le Développement Humain [@arcand2015development]: `indh`
- Voting in the U.S. House of Representatives [@lee2008randomized]: `house`
- STAR dataset [@angrist2008mostly]: `STAR_MHE`

All three data sets are made available as `data.frame` objects.
Using the previously discussed `rdd_data()` function we can transform such a `data.frame` object to an object of class `rdd_data` (which inherits from the `data.frame` object class).

Here we take the data set from the Initiative Nationale pour le Développement Humain (INDH), a development project in Morocco.
The data is included with the package under the name `indh`.

```{r setup, echo=FALSE, message=FALSE}
library(rddtools)
data("indh")
ls()
```

After having loaded the data, we start with inspecting it's structure.

```{r df-str, out.width='\\textwidth'}
str(indh)
```

The `indh` object is a `data.frame` containing 720 observations (representing individuals) of two variables:

- `choice_pg`
- `poverty`

The variable of interest is `choice_pg`, which represent the decision to contribute to a public good or not.
The observations are individuals choosing to contribute or not, these individuals are clustered by the variable `commune` which is the municipal structure at which funding was distributed as part of the INDH project.
The forcing variable is `poverty` which represents the number of households in a commune living below the poverty threshold.
As part of the INDH, commune with a proportion of household below the poverty threshold greater than 30% were allowed to distribute the funding using a **Community Driven Development** scheme.
The cutoff point for our analysis is therefore `30`.

We can now transform the `data.frame` to a special `rdd_data`-class object, inheriting from the `data.frame` class using the `rdd_data()` function.

```{r rdd_data}
rdd_dat_indh <- rdd_data(y=choice_pg,
                         x=poverty,
                         data=indh,
                         cutpoint=30 )
```

The `rdd_data()` can be used using the `data` argument, in which case the function will look for the values of `y` and `x` in this argument (before looking in the `.GlobalEnv`), if this argument is `NULL`, only the `.GlobalEnv` will be scanned.
Additional exogenous variables can be included using the `covar` argument.

The structure is similar to the original `data.frame` object, but contains some additional information.

```{r rdd_data-str}
str(rdd_dat_indh)
```

The `rdd_data` object has the classes `data.frame` and `rdd_data`.
It contains two variables, `y` the explanandum or dependent variable and `x` the explanans or driving variable, which is also our discontinuous variable.
Related to the discontinuous variable is the `attribute` called `cutpoint`, which describes where in the domain of `x` the discontinuity occurs.
The `hasCover` attribute indicates if additional exogenous variables have been included using the `cover` argument to the `rdd_data()` function.


# Analysis

In order to best understand our data, we start with an exploratory data analysis using tables...

```{r summary}
summary(rdd_dat_indh)
```

...and plots.

```{r plot}
plot(rdd_dat_indh)
```

## Parametric Estimation

We can now continue with a standard Regression Discontinuity Design estimation.

```{r reg_para}
reg_para <- rdd_reg_lm(rdd_dat_indh, order=4)
print(reg_para) # uses print.rdd_data
```

and visualising this estimation.

```{r plot-reg_para}
plot(reg_para)
```

## Non-parametric Estimation

In addition to the parametric estimation, we can also perform a non-parametric estimation.

```{r reg_nonpara}
bw_ik <- rdd_bw_ik(rdd_dat_indh)
reg_nonpara <- rdd_reg_np(rdd_object=rdd_dat_indh, bw=bw_ik)
reg_nonpara
```

and visualising the non-parametric estimation.

```{r plot-reg_nonpara}
plot(reg_nonpara)
```

## Sensitivity tests.

In addition to this, sepeveral sensitivity tests for the parametric and non-parametric estimation methods have been implemented.

```{r sensi}
plotSensi(reg_nonpara, from=0.05, to=1, by=0.1)
```

In addition to the sensitivity test, we can also perform various other test such as a placebo test.

# Conclusion and Discussion

The package `rddtools` provides a unified framework for working with Regression Discontinuity Data in `R`.
Functionality already available is several existing packages, such as `rdd` and `KernSmooth` can now easily be utilised using the `rdd_data` framework, as well as several linking functions.

In addition to this, new tools and algorithms have also been implemented
Furthermore, various post-estimation robustness checks are also included in the package.

In addition to the various procedures discussed here,
future packages implementing further RDD functionality can easily leverage the `rdd_data` framework,
which will allow users to quickly access this new functionality through a familiar API.


---
nocite: | 
  @zeilis2006sandwich, @zeileis2006object, @wickham2009ggplot2
...

# References
